---
title: "Week 1 Demo: Interactive Workflow"
---

```{r}
#| echo: false
library(webexercises)
```

## Introduction

This demo showcases the interactive workflow for ECON 2B03.

- **Webexercises**: Fill-in-the-blank, multiple choice, true/false
- **Quarto-Quiz**: Interactive multiple choice with check button
- **Master-Only Content**: Hidden cues for the instructor
- **Placeholders**: Slido & Microsoft Forms

---

# Webexercises Demo

---

## Fill-in-the-blank

Create fill-in-the-blank questions using `fitb()`.

- 2 + 2 is `r fitb(4)`

```{r}
x <- sample(2:8, 1)
```

- The square root of `r x^2` is: `r fitb(x)`
- What is the letter after D? `r fitb("E", ignore_case = TRUE)`

::: {.notes}
Students see instant feedback: green for correct, red for incorrect.
:::

---

## Multiple choice

- Which is a measure of central tendency? `r mcq(c("Variance", answer = "Mean", "Range"))`
- "Never gonna give you up, never gonna: `r mcq(c("let you go", "turn you down", "run away", answer = "let you down"))`"

---

## True or false

- True or False? You can permute values in a vector using `sample()`. `r torf(TRUE)`
- True or False? The mean is always greater than the median. `r torf(FALSE)`

---

## Longer MCQs

**What is a p-value?**

```{r}
opts_p <- c(
   "the probability that the null hypothesis is true",
   answer = "the probability of the observed, or more extreme, data, under the assumption that the null-hypothesis is true",
   "the probability of making an error in your conclusion"
)
```

`r longmcq(opts_p)`

---

# Quarto-Quiz Demo

---

## Quarto-Quiz (Quick Check) {.quiz-question}

Which statement is true?

- The mean is always the median
- [A symmetric distribution has mean = median]{.correct data-explanation="For symmetric distributions, the mean and median coincide."}
- Variance can be negative
- A sample is always the full population

---

## Quarto-Quiz (Multi-Select) {.quiz-question .quiz-multiple}

Select all that apply.

- [The mean is a measure of central tendency]{.correct data-explanation="The mean summarizes central location."}
- [The median is robust to outliers]{.correct data-explanation="The median resists extreme values."}
- Variance can be negative
- [The standard deviation is the square root of variance]{.correct data-explanation="Standard deviation is sqrt(variance), which is non-negative."}

---

## Quarto-Quiz (True/False) {.quiz-question}

The sample mean is an unbiased estimator of the population mean.

- [True]{.correct data-explanation="The expected value of the sample mean equals the population mean."}
- False

---

## Quarto-Quiz (Single Choice) {.quiz-question}

Which term refers to the most frequent value?

- Mean
- Median
- Variance
- [Mode]{.correct data-explanation="The mode is the most frequent value."}

---

## Quarto-Quiz (Ordering) {.quiz-question}

Order these steps for a basic hypothesis test.

1. State the null and alternative hypotheses
2. Choose a significance level
3. Compute the test statistic
4. Compare to a critical value (or compute a p-value)
5. Draw a conclusion

---

## Quarto-Quiz Demo {.quiz-question}

Which probability distribution is symmetric?

- Chi-square
- F-distribution
- [Normal distribution]{.correct data-explanation="The Normal distribution is symmetric around its mean."}
- Exponential

::: {.master-only}
â±ï¸ **Wait 2 minutes** | Teaching Cue: Ask for examples of real-world normal distributions.
:::

---

## Slido Placeholder

::: {.columns}
::: {.column width="50%"}
### Vote Here
*(Slido Embed Placeholder)*
[Link to poll](https://example.com/)
:::

::: {.column width="50%"}
### Live Results
*(Slido Results Placeholder)*
:::
:::

::: {.master-only}
ðŸ”´ **ACTIVATE POLL #1 NOW**
:::

---

## Forms Placeholder

### Participation Credit

**Due Today by 11:59 PM**

[Microsoft Forms Placeholder Link](https://example.com/)

*(QR Code Image Placeholder)*

Enter the secret code from today's lecture: **NORMAL**

::: {.master-only}
Ensure form is open and accepting responses.
:::
