---
title: "Confidence intervals: means and proportions"
part: C
lecture_number: 2
readings: "SZ 7.1, 7.3-7.4"
homework: "SZ 7.3: 1, 3, 5, 7, 9, 13, 15, 19, 21; SZ 7.4: 1, 3, 11, 17"
---

::: {.notes}
2023: did not get to the "minimum sample size" stuff.
:::

# Agenda

## Previously: Part A summary

{{< include ../_summary_A.qmd >}}

## Previously: Part B summary

{{< include ../_summary_B.qmd >}}

## Previously: Part C so far

- confidence intervals are random intervals built around sample estimates
- for means, we already know the large-sample CI when $\sigma$ is known
- we know how to interpret confidence intervals using repeated sampling

## Overview

{{< include ../_readings_homework.qmd >}}

- Why this matters: 
  - in practice, population variability is usually unknown, 
  - many economic questions are about proportions.
- What you will learn:
    1. how to adapt the CI for a population mean when $\sigma$ is unknown
    2. how to construct CIs for population proportions
    3. how sample size affects precision and how to choose a sample for a target margin of error

# Part I: CI for means when $\sigma$ is unknown

## What changes from previous lecture?

- Previously, we assumed the population standard deviation $\sigma$ was known.
- In practice, that is unusual.
- So the practical question is:

> If $\sigma$ is unknown, how do we estimate the standard error of $\bar X$?

. . .

- Answer: we replace $\sigma$ with the sample standard deviation $s$.

## Critical values refresher

We will need critical values that match the confidence level:

::: {.callout-tip title="Three values worth remembering"}
- 90% CI: $z_{\alpha/2} = z_{0.05} \approx 1.645$
- 95% CI: $z_{\alpha/2} = z_{0.025} \approx 1.96$
- 99% CI: $z_{\alpha/2} = z_{0.005} \approx 2.576$
:::

## CI for mu when $\sigma$ is unknown

For $\sigma$ *unknown* and $n$ large, use $s$ instead of $\sigma$:

::: {.callout-important title="Large-sample CI for a population mean ($\sigma$ unknown)"}
The **large-sample $100(1-\alpha)\%$ confidence interval for a population mean  if $\sigma$ is unknown** is $$\bar X \pm z_{\alpha/2} \left(\frac{s}{\sqrt{n}}\right),$$ where a sample is considered large when $n \geq 30$.
:::

## Same logic as previous lecture

- Previously:
  $$\bar X \pm z_{\alpha/2}\left(\frac{\sigma}{\sqrt{n}}\right)$$
- Now:
  $$\bar X \pm z_{\alpha/2}\left(\frac{s}{\sqrt{n}}\right)$$

. . .

- So the main new idea is not a new philosophy.
- The main new idea is: estimate the unknown population variability from the sample.

---

::: {.callout-tip title="Groceries"}
A random sample of 49 students finds:

- sample mean weekly grocery spending: 35 CAD
- sample standard deviation: 14 CAD

Construct a 99% confidence interval for the population mean weekly grocery spending.
Then interpret it.
:::

---

## Worked solution

- Ingredients:
  - $n = 49$, so the large-sample CI is justified
  - $\bar X = 35$
  - $s = 14$
  - for a 99% CI, $z_{0.005} = 2.576$
- Estimated standard error:
  $$\frac{s}{\sqrt{n}}=\frac{14}{\sqrt{49}}=2$$
- Margin of error:
  $$2.576 \times 2 = 5.152$$

```{r}
#| echo: true
35 + 2.576 * 2 * c(-1, 1)
```

. . .

- So the 99% CI is approximately $[29.85, 40.15]$.
- Interpretation:
  - We are 99\% confident that the population mean weekly grocery spending lies between about 29.85 CAD and 40.15 CAD, in the repeated-sampling sense.

## Repeated-sampling reminder

- The interval is random because it depends on the sample.
- The population mean is fixed.
- Confidence refers to the long-run success rate of the method...
- ... not to the probability that this one fixed interval contains $\mu$.

## Exercise: same sample, different confidence levels

:::: {.callout-important .exercise-callout title="Exercise: confidence level and interval width"}
Use the grocery-spending example with $\bar X = 35$, $s = 14$, and $n = 49$.

::: {.panel-tabset}
### Question (2 min)
Without changing the sample, compare the 90% and 99% confidence intervals.

Which one is wider, and why?

### Answers {.answers}
The 99% CI is wider because it uses a larger critical value.

- 90% CI uses $z_{0.05} = 1.645$
- 99% CI uses $z_{0.005} = 2.576$
- The sample and standard error are unchanged
- Higher confidence means a larger margin of error
:::
::::

# Part II: CI for proportions

## Why proportions are special

- Many empirical questions are yes/no questions:
    - employed or not
    - voted or not
    - received a callback or not
- In that setting, the parameter of interest is a population proportion $p$.
- The sample analogue is the sample proportion $\widehat P$.

## CI for p

Large-sample CIs for $p$ are similar to those for $\mu$:

::: {.callout-important title="Large-sample CI for a population proportion"}
The **large-sample $100(1-\alpha)\%$ confidence interval for a population proportion** is
$$\widehat P \pm z_{\alpha/2}\frac{\sqrt{\widehat P(1-\widehat P)}}{\sqrt n}.$$
:::

. . . 

- Use this when the sample is large enough that
  - $n\widehat P \geq 5$
  - $n(1-\widehat P) \geq 5$
- In SZ, the condition is phrased more formally. In class, we will use this simpler rule of thumb.

## Same logic, different standard error

- The structure is still:
    - point estimate $\pm$ critical value times standard error
- What changes is the standard error:
  $$\text{SE}(\widehat P)\approx \sqrt{\frac{\widehat P(1-\widehat P)}{n}}$$

::: {.callout-tip title="Payday loan regulations"}
Suppose a random sample of 826 payday loan borrowers is surveyed about proposed new regulations for payday lenders.

In the sample, 70% support the new regulations.

Construct a 95% confidence interval for $p$, the population proportion of payday borrowers who support these regulations.
:::

---

## Worked solution

- Check the conditions:
  - $\widehat P = 0.70$
  - $n\widehat P = 826 \times 0.70 \approx 578 \geq 5$
  - $n(1-\widehat P) = 826 \times 0.30 \approx 248 \geq 5$
- So the large-sample CI is appropriate.

```{r}
#| echo: true
p_hat <- 0.70
n <- 826
sqrt(p_hat * (1 - p_hat) / n)
p_hat + 1.96 * sqrt(p_hat * (1 - p_hat) / n) * c(-1, 1)
```

. . .

- So the 95% CI is approximately $[0.669, 0.731]$.
- Interpretation:
  - We are 95% confident that the true support rate lies between 66.9% and 73.1%, in the repeated-sampling sense.

---

::: {.callout-tip title="Callback rate"}
Suppose a career centre tracks a random sample of 900 job applications submitted by recent graduates.

Out of those 900 applications, 300 receive a callback.

Construct a 95% confidence interval for $p$, the population callback rate.
:::

---

## Backup example: callback rate solution

- $\widehat P = 300/900 = 1/3$
- The large-sample conditions hold easily.

```{r}
#| echo: true
p_hat <- 300 / 900
p_hat + 1.96 * sqrt(p_hat * (1 - p_hat) / 900) * c(-1, 1)
```

. . .

- So the 95% CI is approximately $[0.30, 0.36]$.

---

## Interpreting the callback example

- Our best estimate of the callback rate is 33.3%.
- A 95% confidence interval for the population callback rate is about 30% to 36%.
- The interval is not centered at 0 or 1 because the observed callback rate is not near those extremes.
- How to interpret $[0.3, 0.36]$? 
  - We are 95% confident that the true callback rate is between 30% and 36%, in the sense that 95% of intervals constructed this way under repeated random sampling would contain the true rate.

## Exercise: six-year graduation rate

:::: {.callout-important .exercise-callout title="Exercise (adapted from SZ 7.3)"}
A university examines 600 randomly selected entering students and finds that 312 graduated within six years.

::: {.panel-tabset}
### Question (3 min)
1. Compute the sample proportion.
2. Construct a 95% CI for the six-year graduation rate.

### Answers {.answers}
- $\widehat P = 312/600 = 0.52$
- Since $600(0.52)$ and $600(0.48)$ are both much larger than 5, the large-sample CI applies.
- The interval is
  $$
  0.52 \pm 1.96 \sqrt{\frac{0.52(0.48)}{600}}
  \approx (0.480,\ 0.560)
  $$
:::
::::

## Exercise: same proportion, different sample sizes

:::: {.callout-important .exercise-callout title="Exercise (adapted from SZ 7.3, Exercise 3)"}
Verify the large-sample conditions, then compare the 90% and 99% CIs for $p$ in each case.

::: {.panel-tabset}
### Question (4 min)
1. $n = 80$, $\widehat P = 0.4$
2. $n = 325$, $\widehat P = 0.4$

### Answers {.answers}
Both satisfy the large-sample conditions.

- For $n=80$, the standard error is about $0.0548$:
  - 90% CI: $0.4 \pm 1.645 \times 0.0548 \approx (0.310,\ 0.490)$
  - 99% CI: $0.4 \pm 2.576 \times 0.0548 \approx (0.259,\ 0.541)$
- For $n=325$, the standard error is about $0.0272$:
  - 90% CI: $0.4 \pm 1.645 \times 0.0272 \approx (0.355,\ 0.445)$
  - 99% CI: $0.4 \pm 2.576 \times 0.0272 \approx (0.330,\ 0.470)$
- In both cases, the 99% CI is wider, and the larger sample gives the narrower interval.
:::
::::

## Takeaway from Parts I and II

- For means and proportions, confidence intervals have the same basic architecture.
- Precision comes from:
    - lower variability
    - larger sample size
    - lower confidence level

# Part III: Minimum sample size

## Turning precision into a design question

So far, we have asked:

- given a sample, how wide is the confidence interval?

Now we ask the reverse question:

- if I want a narrow interval, how large should the sample be?

## Minimum sample size for a mean

::: {.callout-important title="Minimum sample size for estimating a population mean"}
To estimate a population mean to within $E$ units at $100(1-\alpha)\%$ confidence, use

$$n=\frac{(z_{\alpha/2})^2 \sigma^2}{E^2},$$

rounded up to the next integer.
:::

## Where the formula comes from

- Start from the margin of error formula
$$E = z_{\alpha/2}\frac{\sigma}{\sqrt{n}}$$
- Solve for $n$.

. . .

- This is the same logic as before.
- We are just choosing $n$ to hit a target precision level.

## Minimum sample size for a proportion

::: {.callout-important title="Minimum sample size for estimating a population proportion"}
To estimate a population proportion to within $E$ units at $100(1-\alpha)\%$ confidence, use

$$n=\frac{(z_{\alpha/2})^2 \widehat p (1-\widehat p)}{E^2},$$

rounded up to the next integer.
:::

## Choosing $\widehat p$

- Sometimes you have a prior estimate from older data.
- Sometimes you make an informed guess.
- If you want the most conservative sample size, use $\widehat p = 0.5$.

. . .

- Why? Because $\widehat p(1-\widehat p)$ is largest at $0.5$.

## Example: student fees and a football stadium

:::: {.callout-important .exercise-callout title="Exercise (adapted from OpenIntro)"}
A university newspaper wants to estimate the proportion of students who support a \$200 annual fee increase to fund a new football stadium.

::: {.panel-tabset}
### Question (3 min)
How large should the sample be to get a 95% CI with margin of error at most 0.04, if no prior estimate of support is available?

### Answers {.answers}
- Use the conservative choice $\widehat p = 0.50$.
- Then
  $$
  n=\frac{1.96^2(0.5)(0.5)}{0.04^2}\approx 600.25
  $$
- Round up: $n = 601$.
:::
::::

---

## Translating the stadium-fee problem

- Confidence level 95% means $z_{\alpha/2} = 1.96$.
- Target margin of error: $E = 0.04$.
- No prior estimate is available, so use the conservative choice $\widehat p = 0.50$.

---

## Computing the required sample size

```{r}
#| echo: true
1.96^2 * 0.5 * 0.5 / (0.04)^2
```

. . .

- Round up: $n = 601$.
- This is the minimum sample size that guarantees the target margin of error under the conservative assumption.

## Backup example: polling support

::: {.callout-tip title="Optional example"}
A polling firm wants a 90% CI with margin of error 0.5 percentage points for support near 30%.

Estimate the minimum sample size required. Then give the most conservative estimate.
:::

---

## Backup example: polling support solution

- Using $\widehat p = 0.30$:

```{r}
#| echo: true
1.645^2 * 0.3 * 0.7 / (0.005)^2
```

- Rounded up: $n = 22731$.

. . .

- Using the conservative choice $\widehat p = 0.50$:

```{r}
#| echo: true
1.645^2 * 0.5 * 0.5 / (0.005)^2
```

- Rounded up: $n = 27061$.

# Conclusion

{{< include ../_references.qmd >}}

Shafer, Douglas S. and Zhiyi Zhang (2012), "Introductory Statistics", available [via this link.](https://open.umn.edu/opentextbooks/textbooks/135)
